---
title: "Noam's Workbook"
author: "Noam Benkler"
date: "March 28, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(ggplot2)
library(plyr)
library(tidyverse)
library(partykit)
library(rpart)
library(caret)
library(Metrics)
library(dplyr)
library(corrplot)
library(randomForest)
library(readr)
```

#loading data
```{r}
train_dockets <- read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_dockets.csv")
train_termination_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_terminating_motions.csv")
train_other_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_other_motions.csv",header=TRUE)
test_dockets<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_dockets.csv",header=TRUE)
test_termination_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_terminating_motions.csv",header=TRUE)
test_other_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_other_motions.csv",header=TRUE)
```

#Data manipulation
##Wrangling
```{r}
train_dockets <- train_dockets %>% mutate(dismissed =ifelse(outcome == "Dismissed",1,0))
```
##Merging
```{r}
train_other <- full_join(train_dockets, train_other_motions, by = "mudac_id")

train_termination <- full_join(train_dockets, train_termination_motions, by = "mudac_id")

train_full <- full_join(train_dockets, train_other_motions, by = "mudac_id")
train <- full_join(train_full, train_termination_motions, by = "mudac_id")
colnames(train_full)[37:46] <- c("other_motion_type", "other_filing_party", "other_filed_before_joined",
                                 "other_decision", "other_decided_before_jointed", "other_proceeding_percentile",
                                 "termination_motion_type", "termination_filing_party",
                                 "termination_filed_before_joined", "termination_proceeding_percentile")
```

#Questions

#Q4) The MUDAC 2020 challenge is centered around the desire to better understand what influences the final outcome of a case. Recall, the primary outcomes of this investigation are 1) whether or not the case was terminated via a motion to dismiss, 2) whether or not the case was terminated via a motion for summary judgement, and 3) whether or not the case was settled before going to trial

##a) [Level: Beginner/Intermediate] Identify characteristics of a case that appear to be indicative of a case’s outcome
```{r}
#growing trees
train_summary_judgment <- train_dockets %>% select(-c(outcome, settled, dismissed,mudac_id,primarytitle,nos_text))
tree_summary_judgment <- rpart(summary_judgment ~ ., data = train_summary_judgment, method = "class")

train_settled <- train_dockets %>% select(-c(outcome, summary_judgment, dismissed,mudac_id,primarytitle,nos_text))
tree_settled <- rpart(settled ~ ., data = train_settled, method = "class")

train_dismissed <- train_dockets %>% select(-c(outcome, settled, summary_judgment,mudac_id,primarytitle,nos_text))
tree_dismissed <- rpart(dismissed ~ ., data = train_dismissed, method = "class")
```

```{r}
#printing trees
tree_summary_judgment
plot(as.party(tree_summary_judgment), type = "simple")
tree_settled
plot(as.party(tree_settled))
tree_dismissed
plot(as.party(tree_dismissed))
```

```{r}
#growing forrest
rf_summary_judgment <- randomForest(summary_judgment ~ ., data = train_summary_judgment)
rf_settled <- randomForest(settled ~ ., data = train_settled)
rf_dismissed <- randomForest(dismissed ~ ., data = train_dismissed)
```

##b) [Level: Intermediate] Identify any relevant demographic features of a district/court venue that may influence the outcome of a case
```{r}

```


##c)[Level: Intermediate] Investigate any relationships that may exist between the outcome of a case and the various terminating motions that are made
```{r}

```


##d) [Level: Intermediate] Investigate any relationships that may exist between the outcome of a case and the various non-terminating motions that are made
```{r}
tree_summary_judgment_other <- rpart(summary_judgment ~ motion_type + filing_party + filed_before_joined +
                                 decison + decided_before_joined + proceeding_precentile, 
                               data = train_other, method = "class")

tree_settled_other <- rpart(settled ~ motion_type + filing_party + filed_before_joined +
                                 decison + decided_before_joined + proceeding_precentile, 
                               data = train_other, method = "class")

tree_dismissed_other <- rpart(dismissed ~ motion_type + filing_party + filed_before_joined +
                                 decison + decided_before_joined + proceeding_precentile, 
                               data = train_other, method = "class")
```

```{r}
tree_summary_judgment_other
plot(as.party(tree_summary_judgment_other), type = "simple")
tree_settled_other
plot(as.party(tree_settled_other))
tree_dismissed_other
plot(as.party(tree_dismissed_other))
```


##e) [Level: Intermediate] Identify any relationships that may exist between the outcome of a case and whether or not a non-terminating motion is granted/denied










