---
title: "Noam's Workbook"
author: "Noam Benkler"
date: "March 28, 2020"
output: html_document
---
```{r}
data$favor <- mutate(data$favor, 
                     if("filing party" =="Plaintiff" && decision == "Granted"|
                        "filing party" =="Defendant" && decision == "Denied"){"Plaintiff"
                       }else if("filing party"=="Defendant" && decision == "Granted"|
                                "filing party"=="Plaintiff" && decision == "Denied"){"Defendant"
                         }else{NA})
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(ggplot2)
library(plyr)
library(tidyverse)
library(partykit)
library(rpart)
library(caret)
library(Metrics)
library(dplyr)
library(corrplot)
library(randomForest)
library(stargazer)
library(readr)
library(Hmisc)
library(neuralnet)
```

#loading data
```{r}
train_dockets <- read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_dockets.csv")
train_termination_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_terminating_motions.csv")
train_other_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/train_other_motions.csv",header=TRUE)
test_dockets<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_dockets.csv",header=TRUE)
test_termination_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_terminating_motions.csv",header=TRUE)
test_other_motions<-read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/test_other_motions.csv",header=TRUE)
districts <- read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/districts.csv",header=TRUE)
district_fips_code <- read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/district_fips_code.csv",header=TRUE)
demographics_2015 <- read.csv("C:/Users/asus/Dropbox/MUDAC/MUDAC/raw_data/us-census-demographic-data/acs2015_census_tract_data.csv", header = TRUE)
```

#Data manipulation
##Wrangling
```{r}
train_dockets <- train_dockets %>% mutate(dismissed =ifelse(outcome == "Dismissed",1,0)) %>%
  mutate(summary_judgment = as.factor(summary_judgment), settled = as.factor(settled), dismissed =
           as.factor(dismissed))

#demographics
demographics_2015 <- demographics_2015 %>% separate(CensusTract, c("fips_code", "tract"), sep = 5) %>% mutate(fips_code = as.integer(fips_code)) %>% select(-c(tract, State, County))
train_dockets_for_demog <- select(train_dockets, c(district, outcome, summary_judgment, settled, dismissed))
colnames(train_dockets_for_demog)[1] <- "district_number"
district_w_train_dockets <- full_join(train_dockets_for_demog, district_fips_code, by = "district_number")
district_w_demog <- full_join(district_w_train_dockets, demographics_2015, by = c("fips_code"))
district_w_demog_2010_pop <- full_join(district_w_demog, districts, by = c("district_number", "state"))
train_dockets_w_demog <- full_join(district_w_demog_2010_pop, demographics_2015, by = "district_number")
```
##Merging
```{r}
train_other <- full_join(select(train_dockets, c(1,33:36)), train_other_motions, by = "mudac_id")

train_termination <- full_join(select(train_dockets, c(1,33:36)), train_termination_motions, by = "mudac_id")


train_full <- full_join(train_dockets, train_other_motions, by = "mudac_id")
train_full <- full_join(train_full, train_termination_motions, by = "mudac_id")
colnames(train_full)[37:46] <- c("other_motion_type", "other_filing_party", "other_filed_before_joined",
                                 "other_decision", "other_decided_before_jointed", "other_proceeding_percentile",
                                 "termination_motion_type", "termination_filing_party",
                                 "termination_filed_before_joined", "termination_proceeding_percentile")
```

#Questions

#Q4) The MUDAC 2020 challenge is centered around the desire to better understand what influences the final outcome of a case. Recall, the primary outcomes of this investigation are 1) whether or not the case was terminated via a motion to dismiss, 2) whether or not the case was terminated via a motion for summary judgement, and 3) whether or not the case was settled before going to trial

##a) [Level: Beginner/Intermediate] Identify characteristics of a case that appear to be indicative of a case’s outcome
###splitting data
```{r}
#Using LM
train_summary_judgment_lm <- glm(summary_judgment ~ ., data =train_summary_judgment, family = "binomial")
train_settled_lm <- glm(settled ~ ., data =train_settled, family = "binomial")
train_dismissed_lm <- glm(dismissed ~ ., data =train_dismissed, family = "binomial")

stargazer(train_summary_judgment_lm, train_settled_lm, train_dismissed_lm, type = "text")
```



##b) [Level: Intermediate] Identify any relevant demographic features of a district/court venue that may influence the outcome of a case
```{r}
train_summary_judgment_demog <- district_w_demog %>% select(-c(outcome, settled, dismissed))

train_settled_demog <- district_w_demog %>% select(-c(outcome, summary_judgment, dismissed))

train_dismissed_demog <- district_w_demog %>% select(-c(outcome, settled, summary_judgment))

demographics_summary_judgment_lm <- glm(summary_judgment ~ ., data = train_summary_judgment_demog, family = "binomial")
demographics_settled_lm <- glm(settled ~ ., data = train_settled_demog, family = "binomial")
demographics_dismissed_lm <- glm(dismissed ~ ., data = train_dismissed_demog, family = "binomial")
```


##c)[Level: Intermediate] Investigate any relationships that may exist between the outcome of a case and the various terminating motions that are made
```{r}
lm_SJ_train_term <- glm(summary_judgment ~ motion_type, select(train_termination, -c(mudac_id, outcome,settled,dismissed)), family = "binomial")
lm_set_train_term <- glm(settled ~ motion_type, select(train_termination, -c(mudac_id, outcome,summary_judgment,dismissed)), family = "binomial")
lm_dis_train_term <- glm(dismissed ~ motion_type, select(train_termination, -c(mudac_id, outcome,settled,summary_judgment)), family = "binomial")

stargazer(lm_SJ_train_term, lm_set_train_term, lm_dis_train_term, type = "text")
```


##d) [Level: Intermediate] Investigate any relationships that may exist between the outcome of a case and the various non-terminating motions that are made
```{r}
lm_SJ_train_oth <- glm(summary_judgment ~ motion_type, select(train_other, -c(mudac_id, outcome,settled,dismissed)), family = "binomial")
lm_set_train_oth <- glm(settled ~ motion_type, select(train_other, -c(mudac_id, outcome,summary_judgment,dismissed)), family = "binomial")
lm_dis_train_oth <- glm(dismissed ~ motion_type, select(train_other, -c(mudac_id, outcome,settled,summary_judgment)), family = "binomial")

stargazer(lm_SJ_train_oth, lm_set_train_oth, lm_dis_train_oth, type = "text")
```



##e) [Level: Intermediate] Identify any relationships that may exist between the outcome of a case and whether or not a non-terminating motion is granted/denied
```{r}
train_other_decision_twoleveled <- train_other %>% filter(decison == c("Denied", "Denied as Moot", "Granted", "Granted in Part"))
levels(train_other$decison)

condense_decision <- function(x){
  ifelse(x %in% c("Denied","Denied as Moot"), "Denied", "Granted")
}

train_other_decision_twoleveled$decison <- fct_relabel(train_other_decision_twoleveled$decison, condense_decision)

lm_SJ_train_oth <- glm(summary_judgment ~ ., select(train_other_decision_twoleveled, c(summary_judgment, decison)), family = "binomial")
lm_set_train_oth <- glm(settled ~ ., select(train_other_decision_twoleveled, c(settled, decison)), family = "binomial")
lm_dis_train_oth <- glm(dismissed ~ ., select(train_other_decision_twoleveled, c(dismissed, decison)), family = "binomial")

stargazer(lm_SJ_train_oth, lm_set_train_oth, lm_dis_train_oth, type = "text")
```


#Q5) [Level: Advanced] The MUDAC 2020 challenge includes a prediction component. This part of the competition will take place in Kaggle (additional information regarding Kaggle will be provided separately). MUDAC 2020 requires teams to build a predictive model that provides a probability that a case will be closed by a summary judgement and a second predictive model that provides a probability that a case will be closed by a settlement.

##a) Model #1: Obtain a predicted probability that a case will be closed by a summary judgment
###unpruned
```{r}
#Growing tree
train_summary_judgment <- train_dockets %>% select(-c(outcome, settled, dismissed,mudac_id,primarytitle,nos_text))
tree_summary_judgment <- rpart(summary_judgment ~ ., data = train_summary_judgment, method = "class")

#plotting tree
plot(as.party(tree_summary_judgment), type = "simple")
preds_sj <- predict(tree_summary_judgment, train_summary_judgment, type = "prob")
preds_sj[,1]
train_summary_judgment <- train_summary_judgment %>%
  mutate(prob_sj = preds_sj[,2])

preds_settled <- predict(tree_settled, train_settled, type = "prob")
train_settled <- train_settled %>%
  mutate(prob_settled = preds_settled[,2])

train_dockets %>% select(mudac_id) %>% mutate("summary judgment" = preds_sj[,2], settled = preds_settled[,2]) %>% gather("outcome", "probability", -mudac_id) %>% arrange(mudac_id) %>% mutate(Id = paste(mudac_id,outcome)) %>% select(Id, probability)

train_dockets %>% select(mudac_id, outcome) %>% mutate("summary judgment" = preds_sj[,2], settled = preds_settled[,2]) %>% gather("outcome_type", "probability", -c(mudac_id, outcome)) %>% arrange(mudac_id) %>% mutate(Id = paste(mudac_id,outcome_type)) %>% select(Id, probability, outcome)
```


```{r}
#Growing tree
train_summary_judgment <- train_dockets %>% select(-c(outcome, settled, dismissed,mudac_id,primarytitle,nos_text))
tree_summary_judgment <- rpart(summary_judgment ~ ., data = train_summary_judgment, method = "class")

#plotting tree
plot(as.party(tree_summary_judgment), type = "simple")
preds_sj <- predict(tree_summary_judgment, type = "class")

# Adding predictions to the data set
train_summary_judgment <- train_summary_judgment %>%
  mutate(prediction = preds_sj)

# Confusion matrix
train_summary_judgment %>%
  rename(truth = summary_judgment) %>%
  count(truth, prediction) %>%
  spread(key = prediction, value = n)

# accuracy
train_summary_judgment_accuracy <- train_summary_judgment %>% summarise(accuracy = mean(summary_judgment == prediction))
train_summary_judgment_accuracy

# sensitivity
train_summary_judgment_sens <- train_summary_judgment %>%   
  summarise(sensitivity = sum(summary_judgment == "1" & prediction == "1") / 
              sum(summary_judgment == "1"))
train_summary_judgment_sens

# specificity
train_summary_judgment_spec <- train_summary_judgment %>% 
  summarise(specificity = sum(summary_judgment == "0" & prediction == "0") / sum(summary_judgment == "0"))
train_summary_judgment_spec
```

###pruned
```{r}
opt <- which.min(tree_summary_judgment$cptable[,"xerror"])
opt_cp <- tree_summary_judgment$cptable[opt, "CP"]
tree_summary_judgment_prune <- prune(tree_summary_judgment, cp = opt_cp)
tree_summary_judgment_prune
plot(as.party(tree_summary_judgment_prune), type = "simple")

tree_summary_judgment <- rpart(summary_judgment ~ ., data = train_summary_judgment, method = "class", 
                       control = rpart.control(cp = 0.04))
plot(as.party(tree_summary_judgment), type = "simple")

printcp(tree_summary_judgment)
plotcp(tree_summary_judgment)
```

###Random Forrest
```{r}
bagged_summary_judgment_tree <- randomForest(summary_judgment ~ ., data = train_summary_judgment, mtry = 30)
bagged_summary_judgment_tree

rf_summary_judgment_tree <- randomForest(summary_judgment ~ ., data = train_summary_judgment, mtry = sqrt(30))
rf_summary_judgment_tree
```

###Using full training dataset
```{r}

```




```{r}
test %>%
  mutate(
    pred.bag = predict(bagged_tree, newdata = test, type = "class"),
    pred.rf =  predict(rand_forest, newdata = test, type = "class") 
  ) %>%
  summarize(
    bag.error = mean(income != pred.bag),
    rf.error = mean(income != pred.rf)
  )
```

##b) Model #2: Obtain a predicted probability that a case will be closed by a settlement
###unpruned
```{r}
#Growing tree
train_settled <- train_dockets %>% select(-c(outcome, summary_judgment, dismissed,mudac_id,primarytitle,nos_text))
tree_settled <- rpart(settled ~ ., data = train_settled, method = "class")

#plotting tree
plot(as.party(tree_settled), type = "simple")
preds_settled <- predict(tree_settled, type = "class")

# Adding predictions to the data set
train_settled <- train_settled %>%
  mutate(prediction = preds_settled)

# Confusion matrix
train_settled %>%
  rename(truth = settled) %>%
  count(truth, prediction) %>%
  spread(key = prediction, value = n)

# accuracy
train_settled_accuracy <- train_settled %>% summarise(accuracy = mean(settled == prediction))
train_settled_accuracy

# sensitivity
train_settled_sens <- train_settled %>%   
  summarise(sensitivity = sum(settled == "1" & prediction == "1") / 
              sum(settled == "1"))
train_settled_sens

# specificity
train_settled_spec <- train_settled %>% 
  summarise(specificity = sum(settled == "0" & prediction == "0") / sum(settled == "0"))
train_settled_spec
```

###pruned
```{r}
opt <- which.min(tree_settled$cptable[,"xerror"])
opt_cp <- tree_settled$cptable[opt, "CP"]
tree_settled_prune <- prune(tree_settled, cp = opt_cp)
tree_settled_prune
plot(as.party(tree_settled_prune), type = "simple")

tree_settled <- rpart(settled ~ ., data = train_settled, method = "class", 
                       control = rpart.control(cp = 0.04))
plot(as.party(tree_settled), type = "simple")

printcp(tree_settled)
plotcp(tree_settled)
```

###Random Forrest
```{r}
bagged_settled_tree <- randomForest(settled ~ ., data = train_settled, mtry = 30)
bagged_settled_tree

rf_settled_tree <- randomForest(settled ~ ., data = train_settled, mtry = sqrt(30))
rf_settled_tree
```

###Neural Net
```{r}
settled_matrix <- model.matrix( 
  ~ as + Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, 
  data = d 
)

settled_nn <- neuralnet(settled~.,data=train_settled, hidden=5, act.fct = "logistic",
                linear.output = FALSE)
plot(settled_nn)

Predict <- compute(settled_nn,test)
Predict$net.result

prob <- Predict$net.result
pred <- ifelse(prob>0.5, 1, 0)
pred
```

###Running RF on test data
```{r}
test_dockets_cut <- test_dockets %>% select(-c(mudac_id,nos_text)) %>% mutate(summary_judgment = NA) %>% mutate(summary_judgment = as.factor(summary_judgment))
```



```{r}
rf_settled_tree_test <- randomForest(settled ~ ., data = train_settled, mtry = sqrt(30))
rf_summary_judgment_tree_test

train_summary_judgment

test_dockets_cut <- test_dockets %>% select(-c(mudac_id,nos_text))
preds_sj_test <- predict(rf_summary_judgment_tree, test_dockets_cut, type = "prob")
preds_sj_test[,2]
test_summary_judgment <- test_dockets %>% select(mudac_id) %>% mutate("summary judgment" = preds_sj_test[,2])

preds_settled <- predict(tree_settled, train_settled, type = "prob")
train_settled <- train_settled %>%
  mutate(prob_settled = preds_settled[,2])

train_dockets %>% select(mudac_id) %>% mutate("summary judgment" = preds_sj[,2], settled = preds_settled[,2]) %>% gather("outcome", "probability", -mudac_id) %>% arrange(mudac_id) %>% mutate(Id = paste(mudac_id,outcome)) %>% select(Id, probability)

train_dockets %>% select(mudac_id, outcome) %>% mutate("summary judgment" = preds_sj[,2], settled = preds_settled[,2]) %>% gather("outcome_type", "probability", -c(mudac_id, outcome)) %>% arrange(mudac_id) %>% mutate(Id = paste(mudac_id,outcome_type)) %>% select(Id, probability, outcome)

colnames(test_dockets)
colnames(train_dockets)
```


##lets try this again
##a) Model #1: Obtain a predicted probability that a case will be closed by a summary judgment
```{r}
#SJ
train_summary_judgment <- train_dockets %>% select(-c(outcome, settled, dismissed,mudac_id,primarytitle,nos_text))
test_summary_judgment <- test_dockets %>% select(-c(mudac_id, nos_text))

levels(test_summary_judgment$statute) <- levels(train_summary_judgment$statute)
levels(test_summary_judgment$protected_class) <- levels(train_summary_judgment$protected_class)
levels(test_summary_judgment$additional_nos) <- levels(train_summary_judgment$additional_nos)
levels(test_summary_judgment$jury_demand) <- levels(train_summary_judgment$jury_demand)
levels(test_summary_judgment$informa_pauperis) <- levels(train_summary_judgment$informa_pauperis)
levels(test_summary_judgment$class_action) <- levels(train_summary_judgment$class_action)
levels(test_summary_judgment$arbitration_at_filing) <- levels(train_summary_judgment$arbitration_at_filing)
levels(test_summary_judgment$issue_joined) <- levels(train_summary_judgment$issue_joined)
levels(test_summary_judgment$pretrial_conf) <- levels(train_summary_judgment$pretrial_conf)

rf_summary_judgment_tree <- randomForest(summary_judgment ~ ., data = train_summary_judgment, mtry = sqrt(30))
rf_summary_judgment_tree

preds_sj <- predict(rf_summary_judgment_tree, train_summary_judgment, type = "prob")
submission_for_training <- train_dockets %>% select(mudac_id) %>%
  mutate(prob_sj = preds_sj[,2])
submission_for_training

preds_sj_test <- predict(rf_summary_judgment_tree, test_summary_judgment, type = "prob")
submission_for_testing <- test_dockets %>% select(mudac_id) %>%
  mutate(prob_sj_test = preds_sj_test[,2])
submission_for_testing


#Settled
train_settled <- train_dockets %>% select(-c(outcome, summary_judgment, dismissed,mudac_id,primarytitle,nos_text))
test_settled <- test_dockets %>% select(-c(mudac_id, nos_text))

levels(test_settled$statute) <- levels(train_settled$statute)
levels(test_settled$protected_class) <- levels(train_settled$protected_class)
levels(test_settled$additional_nos) <- levels(train_settled$additional_nos)
levels(test_settled$jury_demand) <- levels(train_settled$jury_demand)
levels(test_settled$informa_pauperis) <- levels(train_settled$informa_pauperis)
levels(test_settled$class_action) <- levels(train_settled$class_action)
levels(test_settled$arbitration_at_filing) <- levels(train_settled$arbitration_at_filing)
levels(test_settled$issue_joined) <- levels(train_settled$issue_joined)
levels(test_settled$pretrial_conf) <- levels(train_settled$pretrial_conf)

rf_settled_tree <- randomForest(settled ~ ., data = train_settled, mtry = sqrt(30))
rf_settled_tree

preds_settled <- predict(rf_settled_tree, train_settled, type = "prob")
submission_for_training <- submission_for_training %>%
  mutate(prob_settled = preds_settled[,2])
colnames(submission_for_training) <- c("mudac_id", "SummaryJudgment", "Settled")
submission_for_training %>% gather("outcome", "Probability", -mudac_id) %>% arrange(mudac_id) %>% mutate(id = paste(mudac_id,outcome, sep = "-")) %>% select(id, Probability) %>% arrange(desc(id))

preds_settled_test <- predict(rf_settled_tree, test_settled, type = "prob")
submission_for_testing <- submission_for_testing %>%
  mutate(prob_settled_test = preds_settled_test[,2])
submission_for_testing

colnames(submission_for_testing) <- c("mudac_id", "SummaryJudgment", "Settled")
submission_for_testing  %>% gather("outcome", "Probability", -mudac_id) %>% arrange(mudac_id) %>% mutate(id = paste(mudac_id,outcome, sep = "-")) %>% select(id, Probability) %>% arrange(desc(id))
```


##b) Model #2: Obtain a predicted probability that a case will be closed by a settlement

